
```{r}
library(ggplot2)
library(readr)
library(dplyr)
```

```{r}
aisles = read.csv('aisles.csv')
departments = read.csv('departments.csv')
prior = read.csv('order_products__prior.csv')
train = read.csv('order_products__train.csv')
orders = read.csv('orders.csv')
products = read.csv('products.csv')
```

```{r}
head(aisles)
head(departments)
head(prior)
head(train)
head(orders)
head(products)
```

```{r}
orders$eval_set = as.factor(orders$eval_set)
summary(orders)
```

## Order Level Analysis

**What kind of orders is more likely to contain higher proportion of reorder items?**

```{r}
order_product = rbind(prior, train)
order_reorder = order_product %>%
    group_by(order_id) %>%
    summarise(percentage_reorder = sum(reordered)/n())
order_reorder = merge(orders, order_reorder)
```

```{r}
dow_reorder = order_reorder %>%
    group_by(order_dow) %>%
    summarise(avg_reorder = mean(percentage_reorder))
ggplot(dow_reorder, aes(x = order_dow, y = avg_reorder)) +
    geom_bar(stat = "identity")

# Saturday is a little bit more likely to try new things;
# while Monday people reorder more

# just remember to transform to factors
```

```{r}
time_reorder = order_reorder %>%
    group_by(order_hour_of_day) %>%
    summarise(avg_reorder = mean(percentage_reorder))
ggplot(time_reorder, aes(x = order_hour_of_day, y = avg_reorder)) +
    geom_bar(stat = "identity")
# From 5 to 10, people are more likely to reorder

# set a category as 4-11
```

```{r}
recency_reorder = order_reorder %>%
    group_by(days_since_prior_order) %>%
    summarise(avg_reorder = mean(percentage_reorder))
ggplot(recency_reorder, aes(
        x = days_since_prior_order, y = avg_reorder)) +
    geom_bar(stat = "identity")

# if an order occurs with long recency, then it's less likely to have reorders
# 0 - very high, 1-7 - increase, 8+ - decrease

# add new variables: current_day (0/1), within-1-week (0/1)
```

```{r}
seq_reorder = order_reorder %>%
    group_by(order_number) %>%
    summarise(avg_reorder = mean(percentage_reorder))
ggplot(seq_reorder, aes(
        x = order_number, y = avg_reorder)) +
    geom_bar(stat = "identity")
# the more previous orders, the higher reorders proportion
# but this is not linear
ggplot(seq_reorder, aes(
        x = log(order_number), y = avg_reorder)) +
    geom_bar(stat = "identity")
# kind of better

# log the order_number
```

## Product Level Analysis

**What kind of products are more likely to be reordered?**

```{r}
product_reorder = rbind(prior, train) %>%
    group_by(product_id) %>%
    summarise(percentage_reorder = sum(reordered)/n())
product_reorder = merge(products, product_reorder)
product_reorder = merge(product_reorder, aisles)
product_reorder = merge(product_reorder, departments)
```

```{r}
product_reorder %>%
    arrange(-percentage_reorder)

# Not matched much with dept pattern

# assign product id with the percentile, based on reorder percentage
```

```{r}
product_reorder$reorder_rank = ceiling(100* rank(product_reorder$percentage_reorder) / nrow(product_reorder))

product_rank = select(product_reorder, product_id, reorder_rank)
# but ther are 3 products' id do not appear in this mapping, assign them with default value 50
products = merge(product_rank, products, all.y = TRUE)
products$reorder_rank = ifelse(is.na(products$reorder_rank), 50, products$reorder_rank)

# use this reorder_rank to category products
```



```{r}
dept_reorder = product_reorder %>%
    group_by(department) %>%
    summarise(avg_reorder = mean(percentage_reorder))
ggplot(dept_reorder, aes(
        x = department, y = avg_reorder)) +
    geom_bar(stat = "identity") +
    coord_flip()

# very obvious differnces
```

```{r}
aisle_reorder = product_reorder %>%
    group_by(aisle) %>%
    summarise(avg_reorder = mean(percentage_reorder))
ggplot(aisle_reorder, aes(
        x = aisle, y = avg_reorder)) +
    geom_bar(stat = "identity") +
    coord_flip()

# very obvious differnces
```

## Dataset Preparation


only include those user-id in train set to train the model.
features - 
    relate to product_id: product_rank, dept, aisle
    relate to user_id: order_num, order_dow (factor), order_hour_of_day (4-11 or not), days_since_prior_order(add same_day, within_one_week)
    relate to past history: prior_count

create a unique mapping: user_id to last order_num, order_dow, ...
(simply filter orders by train+test)
merge it with the prior_count to create a train+test set with eval_set label remained. 
merge the product info to it.
then just split into two. all those in test set should have NA for reorder column

```{r}
product_train = filter(merged, eval_set == "train")
product_prior = filter(merged, eval_set == "prior")

product_train = product_train[product_train$reordered == 1,]


prior_count2 = merge(prior_count, product_train,
                     by.x = c("user_id", "product_id"),
                     by.y = c("user_id", "product_id"),
                     all.x = TRUE)
prior_count2 = select(prior_count2, user_id, product_id, count, reordered)

train_test = filter(orders, eval_set != "prior")
prior_count2 = merge(prior_count2, train_test,
                     by.x = "user_id", by.y = "user_id")
prior_count2 = merge(prior_count2, products,
                     by.x = "product_id", by.y = "product_id",
                     all.x = TRUE)

```

```{r}
prior_count2$product_name = NULL
prior_count2$order_dow = as.factor(prior_count2$order_dow)
prior_count2$peak_hour = ifelse(prior_count2$order_hour_of_day >= 4 & prior_count2$order_hour_of_day<= 11, 1, 0)
prior_count2$same_day = 
    ifelse(prior_count2$days_since_prior_order == 0, 1, 0)
prior_count2$within_1week = 
    ifelse(prior_count2$days_since_prior_order <= 7 & 
               prior_count2$days_since_prior_order > 0, 1, 0)

prior_count2$reorder_freq = prior_count2$count / prior_count2$order_number
```

```{r}
prior_count2$order_number_log = log(prior_count2$order_number)

prior_count2[is.na(prior_count2$reordered),]$reordered = 0

dept_merged = merge(merged, products, 
                    by.x = "product_id", by.y = "product_id")

dept_reorder = dept_merged %>%
    group_by(department_id) %>%
    summarise(percentage_reorder = sum(reordered) / n())

dept_reorder$reorder_rank = 
    nrow(dept_reorder) - rank(dept_reorder$percentage_reorder)

aisle_reorder = dept_merged %>%
    group_by(aisle_id) %>%
    summarise(percentage_reorder = sum(reordered) / n())

aisle_reorder$reorder_rank = nrow(aisle_reorder) - rank(aisle_reorder$percentage_reorder)

prior_count2 = merge(prior_count2, dept_reorder,
                     by.x = "department_id", 
                     by.y = "department_id")

prior_count2 = merge(prior_count2, aisle_reorder,
                     by.x = "aisle_id", 
                     by.y = "aisle_id")


prior_count2$reorder_rank_p = prior_count2$reorder_rank.x
prior_count2$reorder_rank_d = prior_count2$reorder_rank.y
prior_count2$reorder_rank_a = prior_count2$reorder_rank
prior_count2$aisle_id = NULL
prior_count2$department_id = NULL
prior_count2$reorder_rank.x = NULL
prior_count2$reorder_rank.y = NULL
prior_count2$reorder_rank = NULL
prior_count2$percentage_reorder.x = NULL
prior_count2$percentage_reorder.y = NULL

```


```{r}
train = filter(prior_count2, eval_set == "train")
test = filter(prior_count2, eval_set == "test")
write.csv(train, "train.csv")
write.csv(test, "test.csv")
```
